Correlation among smart grid actors

### 2015-06-25 Article.tex
some updates

Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences. One obvious condition for stability revolves around sustaining at any time a power generation that meets the demand. If one is larger than the other, the system deviates from its synchronous state. If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures. With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption. Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities. Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid. It is often assumed that the granularity of these prices is meant to increase in the future as a way to reverberate the production conditions on the end users. The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market. Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants. These kind of settings are already being used at the transport level with large renewable plants in the production portfolio, and they appear as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.

The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed. This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks. Nodes that were simply pure loads yesterday could behave tomorow as generators or loads depending on time. On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future. These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...), which predictions can be difficult and uncertain. Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition. The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.

@@

In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity. More precisely, we consider a set of agents that can be seen as aggregators of renewable distributed energy ressources (DER) and electrical loads. The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid. Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).

As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want. Actually the architecture and control of such a distributed grid creates numerous interdisciplinary challenges. In this paper, the envisaged solution for organizing such a multi-producers system consists in extending the kind of market environement discussed in the previous section. Participating aggregators estimate a production capacity for an upcoming period of time and announce it on the market. If a contract between the grid and a participant is bought, the latter commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.

@@

Forming coalitions/aggregations is an envisaged solution to both the number of entities sollicitating the operator, and their high variabilities. Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side, since an aggregator can be considered as a single node by the level above it. On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio. One thus expect a more stable and predictable energy production for a coalition than for a single agent. Nevertheless, all coalitions are obviously not equally stable, which means that special attention should be paid to the aggregation step.

@@

We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions. Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell. More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions. We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market. Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide. Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or intentional false information) will be exposed to financial penalities.

@@

\subsection{Algorithm}

@@

The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities. Nevertheless, as it comprises mainly of a greedy optimization based on local improvements, the probability that the algorithm finds the global optimal coalitions set is very low. Actually, it is not obvious that a strict optimum exists at all. Besides, there is, to our knowledge, no state of the art algorithm that aggregate uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems).

In order to have an idea about the algorithm's quality, we compare its results with :
\begin{itemize}
\item \textbf{Random :} This algorithm is basically a random search over the coalition structures space. It analyses a given number of structures and returns the best that it has encountered so far. See algorithm \ref{alg:algo2}.
\item \textbf{Correlated :} This is the complete opposite of our algorithm. It uses the correlation graph $ G_{1} $ and performs a community detection. The resulting coalitions have thus very high internal correlations. We thus expect this algorithm to perform very bad compared to the others. See algorithm \ref{alg:algo3}.
\end{itemize}

@@

\begin{figure}
\includegraphics[scale=.6]{./figures/coalitions/coalitions.pdf}
\caption{Coalitions formed in the (contract value, volatility) space. The color map indicates qualities of portions of the plane. The closer to red the better (high contract values with small volatility). On the opposite, blue areas shows poor quality (small contract values with high volatility). Blue dots stand for the un-correlated coalitions, red triangles for random search, and green squares for correlated coalitions.}
\label{fig:coalitions}
\end{figure}

@@

>>>>

Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences. One obvious condition for stability revolves around sustaining at any time a power generation that meets the demand. If one is larger than the other, the system deviates from its synchronous state. If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures. With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption. Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities. Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid. It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users. The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market. Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants. These kind of settings are already being used at the transport level of current power grid with large renewable plants in the production portfolio. Using electricity markets appears thus as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.

The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed. This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks. Nodes that were simply pure loads yesterday could behave tomorow as generators or loads. On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future. These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...). Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition. The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.

@@

In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity. More precisely, we consider a set of agents that can be seen as aggregators of renewable distributed energy ressources (DER) and electrical loads. The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid. Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper). The primary objective of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function. While the latters intend to maximize their benefits with higer production contracts, the operator primary concerns are linked to the quality and stability of the injections. The coalitions that we wish to form should thus be both stable and productive as to satisfy both parts. This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation relationships come into play.

As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state forbid any anarchic system where prosumers inject power without proper control. Actually the architecture and control of such a distributed grid creates numerous interdisciplinary challenges. In this paper, the envisaged solution for organizing such a multi-producers system consists in extending the kind of market environement discussed in the previous section. Participating aggregators estimate a production capacity for an upcoming period of time and announce it on the market. If a contract between the grid and a participant is bought, the latter commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.

@@

Forming coalitions/aggregations is an envisaged solution to both the number of entities sollicitating the operator, and their high variabilities. Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side, since an aggregator can be considered as a single node by the level above it. On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio. One thus expect a more stable and predictable energy production for a coalition than for a single agent. Nevertheless, all coalitions are obviously not equally stable or productive, which means that special attention should be paid to the aggregation step.

@@

We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions. Understanding the correlation reliationships among the agents can thus give an indication about what coalitions to form and how much they should sell. More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions. We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market. Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide. Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or intentional false information) will be exposed to financial penalities.

@@

@@

The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities. As it comprises mainly of a greedy optimization based on local improvements, there is no guarantee that the algorithm finds the global optimum. Since there is, to our knowledge, no state of the art algorithm that aggregates uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems), we compare the results with :

\begin{itemize}
\item \textbf{Random :} This algorithm is basically a random search over the coalition structures space. It analyses a given number of structures and returns the best that it has encountered so far. See algorithm \ref{alg:algo2}.
\item \textbf{Correlated :} This is the complete opposite of our algorithm. It uses the correlation graph $ G_{1} $ and performs a community detection. The resulting coalitions have thus very high internal correlations. We thus expect this algorithm to perform very bad compared to the others. See algorithm \ref{alg:algo3}.
\end{itemize}

@@

\begin{figure}
\includegraphics[scale=.5]{./figures/coalitions/coalitions6.pdf}
\caption{Coalitions formed in the (contract value, volatility) space. The color map indicates qualities of portions of the plane. The closer to red the better (high contract values with small volatility). On the opposite, blue areas shows poor quality (small contract values with high volatility). Blue dots stand for the un-correlated coalitions, red triangles for random search, and green squares for correlated coalitions.}
\label{fig:coalitions}
\end{figure}

@@

\appendix
\subsection{Algorithms}

\subsection{Net production series}

Data were collected from [] (similar data can be found at [] for the United States). The variables used in the simulation are :

\begin{itemize}
\item Average wind speed (in $m.s^{-1}$)
\item Nebulosity (integer in $[0,8] $)
\item Temperature (in degree celsius)
\end{itemize}

Power curves are functions that, for a given type of generator, map the input to the output power. For windturbines and solar arrays these functions are well studied and approximations have been proposed. For the wind turbines, the power curve is thus specified by 4 values :

\begin{itemize}
\item Cut-in-speed : The wind speed at which the turbine first starts to rotate and generates power.
\item Rated-output-power : The maximum power that the turbine can generate.
\item Rated-output-speed : The wind speed at which the turbine attains its rated output power.
\item Cut-out-speed : The speed at which the turbine is turned off as not to damage the rotor.
\end{itemize}

The most interesting part is the increase of output power when the wind speed is in the cut-in-speed rated-output-speed range. Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the litterature.

The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data. As we mainly collected nebulosity series, we used the Helios model described in []. This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps. As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see [] for more details) :

\begin{equation}
\left\{ \begin{array}{lll}
			\Psi{real}(t) = \Psi_{perfect}(t) \eta(t) \\
			\eta(t) = 1-0.75 \left( \dfrac{N(t)}{8} \right)^{3.4}
\end{array} \right.
\end{equation}

Where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.

Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :

\begin{equation}
\mathcal{F}_{PV}(\Psi{real}(t)) =  S_{PV} \Psi{real}(t)  e_{PV}
\end{equation}

Where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency. The very simple form of this power curve is due to some simplifications in order not to overload the model. For instance, it does not take into account angles and orientations degradations. These could be incorporated if needed by changing the power curve in the simulations.

data, helios model, power curves, scema
